A properly labelled proof system for intuitionistic modal logics

### 2019-09-10 caleidoscope/caleidoscope.tex
done section 3

Now we can present the inference rules for \emph{system $\labIKp$} for the logic $\IK$.
We obtained this system, shown in Figure~\ref{fig:labIKp}, as follows.
Our starting point was the multiple-conlusion nested sequent system \emph{\`a la} Maehara (as presented in~\cite{kuz:str:maehara}), which can straightforwardly translated into the labelled setting, and which yields the rules $\llabrn\bot$ \todo{$\TOP$ yes/no}, $\llabrn\AND$, $\rlabrn\AND$, $\llabrn\OR$, $\rlabrn\OR$, $\llabrn\DIA$, and $\rlabrn\DIA$ as we show them in Figure~\ref{fig:labIKp}. However, this translation also yields the rules $\rn{id'}$, $\llabrn\IMP'$, and~$\llabrn\BOX'$:
\begin{equation}
  \label{eq:nomon}
  \vlinf{\rn{id}'}{}{\B, \Left, \labels{x}{a} \SEQ \Right, \labels{x}{a} }{}
  \quad
  \vliinf{\llabrn\IMP'}{}{\B,\Left, \labels{x}{A} \IMP B \SEQ \Right}{\B,\labels{x}{A \IMP B}, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left, \labels{x}{B} \SEQ \Right}
  \quad
  \vlinf{\llabrn\BOX}{}{\B, xRz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, xRz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}
\end{equation}
that are not sufficient for a complete system, and we will see below why. Before, let us first look at the rules $\rlabrn\IMP$ and $\rlabrn\BOX$. In the  multiple-conlusion nested sequent system of~\cite{kuz:str:maehara}, these are the two rules that force single-conclusion. This is the reason why in our system we need the $\le$ in the relational atoms; we need to go to the future of the current world. 
In the Kripke-semantics in~\eqref{eq:kripke} the two connectives $\IMP$ and $\BOX$ are the ones that need the future-relation $\le$. This relation is reflexive and transitive. In order to capture that in the proof system, we need to add the rules $\rn{refl}$ and $\rn{trans}$. Finally, in the semantics, the two relations $\rel$ and $\le$ are strongly connected through the two conditions $F_1$ and $F_2$ \todo{proper typography}. These need to be reflected in the proof system, which is done by the two rules $\rn{F_1}$ and $\rn{F_2}$.

\todo{monotonicity}

>>>>

Now we can present the inference rules for \emph{system $\labIKp$} for the logic $\IK$.
We obtained this system, shown in Figure~\ref{fig:labIKp}, as follows.
Our starting point was the multiple-conlusion nested sequent system \emph{\`a la} Maehara (as presented in~\cite{kuz:str:maehara}), which can straightforwardly translated into the labelled setting, and which yields the rules $\llabrn\bot$ \todo{$\TOP$ yes/no}, $\llabrn\AND$, $\rlabrn\AND$, $\llabrn\OR$, $\rlabrn\OR$, $\llabrn\DIA$, and $\rlabrn\DIA$ as we show them in Figure~\ref{fig:labIKp}. However, this translation also yields the rules $\rn{id'}$, $\llabrn\IMP'$, and~$\llabrn\BOX'$:
\begin{equation}
  \label{eq:nomon}
  \vlinf{\rn{id}'}{}{\B, \Left, \labels{x}{a} \SEQ \Right, \labels{x}{a} }{}
  \quad
  \vliinf{\llabrn\IMP'}{}{\B,\Left, \labels{x}{A} \IMP B \SEQ \Right}{\B,\labels{x}{A \IMP B}, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left, \labels{x}{B} \SEQ \Right}
  \quad
  \vlinf{\llabrn\BOX}{}{\B, xRz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, xRz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}
\end{equation}
that are not sufficient for a complete system, and we will see below
why. Before, let us first look at the rules $\rlabrn\IMP$ and
$\rlabrn\BOX$. In the multiple-conlusion nested sequent system
of~\cite{kuz:str:maehara}, these are the two rules that force
single-conclusion. This is the reason why in our system we need the
$\le$ in the relational atoms; we need to go to the future of the
current world.  In the Kripke-semantics in~\eqref{eq:kripke} the two
connectives $\IMP$ and $\BOX$ are the ones that need the
future-relation $\le$. This relation is reflexive and transitive. In
order to capture that in the proof system, we need to add the rules
$\rn{refl}$ and $\rn{trans}$.

Finally, in the semantics, the two relations $\rel$ and $\le$ are
strongly connected through the two conditions $F_1$ and $F_2$
\todo{proper typography}. These need to be reflected in the proof
system, which is done by the two rules $\rn{F_1}$ and $\rn{F_2}$.
%
However, these two rule create new labels, and in order to be
complete, the system needs the \emph{monotonicity} rule:
\begin{equation}
  \label{eq:mono}
  \vlinf{\rn{mon}}{}{\lseq{\B, x \le x'}{\Left, \labels{x}{A}}\Right}{
    \lseq{\B, x \le x'}{\Left, \labels{x}{A}, \labels{x'}{A}}\Right}
\end{equation}
Since this rule is a form of contraction, it would cause the same
problems as contraction in a cut elimination proof. Hence, it is
preferable to have a system in which this rule is admissible. This is
the reason why we have monotonicity incorporated in the rules
$\rn{id}$, $\llabrn\IMP$ and $\llabrn\BOX$ in Figure~\ref{fig:labIKp},
instead of using the rules in~\eqref{eq:nomon}.

\begin{theorem}
  \label{thm:mon-adm}
  The rule $\rn{mon}$ is admissible for system $\labIKp$.
\end{theorem}

For proving this theorem, we need a series of auxiliary lemmas.

\todo{}

\begin{proof}[Proof of Theorem~\ref{thm:mon-adm}]
  \todo{}
\end{proof}

In the following sections, we will show that the system $\labIKp$ is sound and complete. For the completeness proof we proceed via cut elimination. The cut rule has the following shape:
\begin{equation}
  \label{eq:cut}
  \vliiinf{\labrn{cut}}{}{
    \lseq\B\Left\Right}{
    \lseq\B\Left{\Right,\labels{z}{C}}}{}{
    \lseq\B{\Left,\labels{z}{C}}{\Right}}
\end{equation}
Then we can summarize soundness and completeness, and cut
admissibility of $\labIKp$ in the following Theorem:

\begin{theorem}\label{thm:cutfree-compl}
	%	Let $\CC$ be a set of geometric frame properties as in~\eqref{eq:cla-geometric} and $\labbrn{\CC}$ be the corresponding set of rules following schema~\eqref{eq:modal-grs}.
	%
	For any formula $A$, the following are equivalent.
	%
	\begin{enumerate}
		\item\label{i} $A$ is a theorem of $\IK$. 
		%
		\item\label{ii} $A$ is provable in $\labIKp +\labrn{cut}$.
		\item\label{iii} $A$ is provable in $\labIKp$.
		%
		\item\label{iv} $A$ is valid in every birelational frame. %satisfying the properties in $\CC$.
	\end{enumerate}
\end{theorem}

The proof of this theorem is the topic of the following sections. The
equivalence of \ref{i} and \ref{iv} has already been stated in
Theorem~\ref{thm:plotkin}~\cite{Fischer,Plotkin}. The implication
\ref{i}$\implies$\ref{ii} is shown in Section~\ref{sec:completeness},
the implication~\ref{ii}$\implies$\ref{iii} is shown in
Section~\ref{sec:cut-elim}, and finally, the implication
\ref{iii}$\implies$\ref{iv} is shown in Section~\ref{sec:soundness}.

\section{Soundness}\label{sec:soundness}

\section{Completness}\label{sec:completeness}

\section{Cut Admissibility}\label{sec:cut-elim}
